\section{Introduction}
Power efficiency has become a critical design factor with the continual demands of emerging applications including machine learning and scientific computation. As alternatives to general purpose CPUs and GPUs , customized hardware accelerators~\cite{extensor, fpdnn, bigdata_fpga} like ASICs and FPGAs provide good energy efficiency and performance. However, hardware design is still a big challenge. Hardware description languages (HDLs) including Verilog~\cite{verilog} and VHDL~\cite{vhdl} are frequently used in industry for hardware design. These HDLs adopt a low level of abstraction including wires, registers, and gates, also known as register transfer level (RTL). However, HDLs' productivity is seriously hampered by this low level of abstraction, which makes it difficult for hardware designers.

Both high-level synthesis (HLS) and hardware generators have been proposed to improve the productivity of hardware design. HLS is a general method for automatically generating hardware from a behavioral description written in high-level languages such as C, C++, or OpenCL~\cite{Dynamatic_Josipovi, uIR_Sharifian, flexcl}. HLS offers a promising way to design hardware in a high-level abstraction, which provides opportunities for software engineers without hardware experience to design hardware quickly. However, it can lead to bad performance and high resource utilization due to the lack of domain knowledge about the computation and architecture~\cite{hls_liang}. On the other hand, hardware generator makes use of domain knowledge to improve productivity. These generators often focus on a specific domain, such as streaming~\cite{aetherling,parallel_pattern} and spatial accelerators~\cite{tensorlib,ems}. With the domain knowledge of the target applications, these approaches use specific languages and synthesis flows to generate hardware with good performance. Although these approaches provide high level abstraction, they can only support a specific class of applications.

%Existing HLS tools~\cite{VivadoHls, legup, bambu, soda} don't take readability and flexibility into account, which makes the generating process opaque and cryptic.} 

% \hl{this paragraph gives an overview of the systhesis flow for each technique.} And HLS compilers also implement a series of transformations that makes up an end-to-end flow from high-level language to HDL program. 
Furthermore, different synthesis tools often adopt different synthesis methodologies. HLS compilers convert the high-level description to hardware implementation in three steps: allocation, scheduling, and binding. The scheduling step, which determines the cycles for each operation, can be implemented using different algorithms including static, dynamic, and hybrid scheduling~\cite{Dynamatic_Josipovi, cheng_dss}. Hardware generators perform sophisticated architectural transformations to improve performance and resource consumption. In addition to the domain specific optimizations, hardware generators are often guided with specific hardware templates to generate hardware implementation for different applications~\cite{tensorlib, aetherling}. %\hl{As for hardware generators, a domain-specific language (DSL) is often proposed to describe computation and achieve inner optimizations. Each DSL compiler must implement a complete flow that can generate HDL in order to convert DSL into hardware, resulting in numerous redundant implementations.}

% \hl{this paragraph should introduce the flows in details such as IRs.}
% \hl{Adopting fully separate synthesis flows makes it difficult to combine different methods. HLS offers a convenient way to design hardware, but it can only achieve conservative parallelism due to the omission of computation characteristics. In addition, hardware generators can provide efficient hardware with domain-specific optimizations, making it appropriate for stressful circumstances. The combination of different methodologies is a superior option for achieving good performance while maintaining enough productivity. However, programmers can only connect the generated hardware at the RTL level to combine different methods.}
% \hl{this is not a good reason.}

%The spatial accelerator including PE and interconnection is finally generated by different hardware templates.

Both synthesis flows require substantial engineering efforts due to RTL’s weak expressivity and low level of abstraction. HLS tools take high-level languages as inputs and use the compiler infrastructure. For example, Vitis HLS~\cite{VivadoHls} adopts LLVM IR~\cite{lattner2004llvm} as its internal representation, which is a widely used software compilation IR. After that, additional passes are needed to convert to hardware semantics. As for hardware generators, domain specific languages (DSL) and optimizations are often used for description and optimization. For example, Tensorlib~\cite{tensorlib} selects tensor algebra as its representation, and maps it into the hardware PE array using space time transformation.  However, these different methodologies share some similarities in the intermediate representations, such as control logic and hardware generation from a high-level abstraction. Therefore, it is possible to have a unified IR for different synthesis approaches so that users can easily design new hardware synthesis techniques based on the same infrastructure and explore different methodologies.


% \hl{here, we should talk about how difficult is for each approach, it is better use detailed IRs and tool chains as examples, and emphasize that they are different}. HLS tools take high-level languages as input and gain enough optimizations from the compiler infrastructure. Therefore, the intermediate representation (IR) of the language is also adopted by these HLS tools. As for hardware generators, a domain-specific language (DSL) is frequently proposed to describe computation and achieve optimizations. All these synthesis methodologies need a complete flow that can generate synthesizable hardware, resulting in numerous redundant implementations. \hl{However, these flows share some similarities in the intermediate representations, then give some insights. Then, it will be great to have unified IR for different synthesis approaches so that}.

%To avoid duplicated implementations and to explore better combinations of multiple approaches, a flexible intermediate representation is needed.

In this paper, we propose Hector, a two-level IR providing a unified description for different hardware synthesis methodologies with expressivity and flexibility. The high-level IR (TOpological Representation) binds computations with a control graph annotated with timing information, while the low-level IR (Hierarchical Elastic Componenent) provides a concise way to describe various hardware components and elastic interconnections among them using customizable primitives. Both IRs provide a uniform representation of the control logic with various manners, but at different abstraction levels. The IRs in Hector are converted to synthesizable RTL programs through a series of transformations including time graph transformation, lowering pass, and RTL generation. The two-level IR and all the transformations are built on the multi-level compiler infrastructure (MLIR)~\cite{mlir}. 

The main contributions of this work are as follows:
\begin{itemize}
\item We propose Hector, a unified IR and synthesis framework supporting multiple hardware synthesis methodologies including high-level synthesis and hardware generator. 
\item We propose a two-level IR that is general enough for different hardware synthesis methdologies. 
\item We implement a compiler that lowers the IRs to synthesizable RTL through a series of analysis  and optimizations.

%Hector provides the ability to design hardware with multiple synthesis approaches. 

%\item We establish several transformation passes to lower from \tor to \hec considering hardware with different behaviors.
% \item We use Hector to implement three hardware synthesis methods including high-level synthesis, systolic array generator and an accelerator for sparse linear algebra, which further demonstrates Hector's convenience and extensiblity.
\end{itemize}

To demonstrate the expressivity and versatility, we implement three synthesis approaches based on Hector: a high-level synthesis (HLS) tool, a systolic array generator, and an accelerator component for sparse linear algebra. The experimental results show that the hardware generated by Hector's hybrid HLS scheduling improves the performance by 29$\%$ on average compared to that generated by the state-of-the-art HLS tool. The other two approaches outperform HLS
implementations in performance and productivity. \textit{Hector is open source at GitHub (\url{https://github.com/pku-liang/Hector})}. 
% \hl{please give the detailed experiment data.}